---
title: "HW2" 
Author:Jingyan Xiong 
---
install.packages("dplyr")
install.packages("tidyr")
install.packages("reshape2")
install.packages("data.table")
install.packages("xtable")
install.packages("bayesm")
install.packages("matrixStats")
library(dplyr)
library(tidyr)
library(reshape2)
library(data.table)
library(xtable)
library(bayesm)
library(matrixStats)

## Exercise 1 Data Description
```{r}
data("margarine")
choicePrice<-margarine$choicePrice
demos<-margarine$demos

# Average in product characteristics (price)
allprice<-choicePrice[,3:12]
average=apply(allprice,2,mean)
print(average)

#Dispersion in product characteristics (price)
dispersion <- diag(var(allprice))
print(dispersion)


# Market share (choice frequency) and market share by product characteristics
choicefreq <- data.frame(table(choicePrice$choice)) 
x <- choicefreq$Freq
y <- x/nrow(choicefreq)
marketshare <- data.frame(y)
colnames(marketshare) <- c("marketshare")
rownames(marketshare) <- c("PPk_Stk", "PBB_Stk", "PFl_Stk",  "PHse_Stk", "PGen_Stk", "PImp_Stk", "PSS_Tub",  "PPk_Tub",  "PFl_Tub",  "PHse_Tub")
print(marketshare)

#Illustrate the mapping between observed attributes and choices.
mm = merge(choicePrice, demos, by = "hhid", all.x = TRUE)
ms = cbind(temp[,1:2],temp[,13:19])
View(ms)
```



## Exercise 2 First Model
```{r}
x=nrow(choicePrice)
y=ncol(choicePrice[,3:12])
Y=matrix(0,x,y)
for(i in 1:y){
  for(j in 2:x){
    if(choicePrice$choice[j]==i){
      Y[j,i]=1
    }
  }
}
Y[1,1]=1

clogit=function(beta){
 intercept=matrix(rep(beta[1:y-1],each=x),nrow=x,ncol=y-1)
 XB=choicePrice[,3:12]*beta[y]
 XB=cbind(0,intercept)+XB
 eXB=exp(XB)
 teXB=rowSums(eXB)
 prob=eXB/teXB
 like=sum(Y*log(prob))
 return(-like)
}
set.seed(100)
param1=runif(10,-1,1)
model1=optim(param1,clogit,method="BFGS")
model1$par
```


## Exercise 3 Second Model
```{r}
merge1 <- merge(demos[,c(1:2)],choicePrice[,c(1:2)], by="hhid")
income <- matrix(rep(merge1[,2],10),ncol=10)
MultLogit <- function(theta, X, Y){
  alpha = sapply(c(0,theta[10:18]),rep,4470)
  beta = sapply(c(0,theta[1:9]) ,rep,4470)
  temp = prop.table(exp(X*beta + alpha),1)
  Y = -sum(k1_10*log(temp))
  return(Y)
}

optim(c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0), MultLogit, X=income, method = "BFGS")$par
```

## Exercise 4 Marginal Effects
```{r}
a <- m1[2:10]
a <- sapply(a,rep,4470)
a2 <- matrix(c(0),nrow = 4470,ncol = 1)
a <- cbind(a2,a)

price <- as.data.frame(choicePrice[,3:12])
prt = NULL
for(i in 1:10){
  prt <- cbind(prt, price[,i] - price[,1])
}
x <- prt

vij <- x*m1[1]+a
total <- exp(vij[,1])+exp(vij[,2])+exp(vij[,3])+exp(vij[,4])+exp(vij[,5])+exp(vij[,6])+exp(vij[,7])+exp(vij[,8])+exp(vij[,9])+exp(vij[,10])
pij <- exp(vij)/total

dataset <- matrix(c(0),ncol=10,nrow=4470)
dataset <- cbind(as.matrix(choicePrice[,2]),dataset)
for(i in 1:4470){
  for(r in 1:10){
    dataset[i,r+1] <- ifelse(dataset[i,1]==r,1,0)
  }
}
dij <- dataset[,2:11]

pij_1 <- as.matrix(cbind(pij[,1],pij[,1],pij[,1],pij[,1],pij[,1],pij[,1],pij[,1],pij[,1],pij[,1],pij[,1]))
me_1 <- pij_1*(dij-pij)*m1[1,]
me_1 <- apply(me_1,2,mean)
pij_2 <- cbind(pij[,2],pij[,2],pij[,2],pij[,2],pij[,2],pij[,2],pij[,2],pij[,2],pij[,2],pij[,2])
me_2 <- pij_2*(dij-pij)*m1[1,]
me_2 <- apply(me_2,2,mean)
pij_3 <- cbind(pij[,3],pij[,3],pij[,3],pij[,3],pij[,3],pij[,3],pij[,3],pij[,3],pij[,3],pij[,3])
me_3 <- pij_3*(dij-pij)*m1[1,]
me_3 <- apply(me_3,2,mean)
pij_4 <- cbind(pij[,4],pij[,4],pij[,4],pij[,4],pij[,4],pij[,4],pij[,4],pij[,4],pij[,4],pij[,4])
me_4 <- pij_4*(dij-pij)*m1[1,]
me_4 <- apply(me_4,2,mean)
pij_5 <- cbind(pij[,5],pij[,5],pij[,5],pij[,5],pij[,5],pij[,5],pij[,5],pij[,5],pij[,5],pij[,5])
me_5 <- pij_5*(dij-pij)*m1[1,]
me_5 <- apply(me_5,2,mean)
pij_6 <- cbind(pij[,6],pij[,6],pij[,6],pij[,6],pij[,6],pij[,6],pij[,6],pij[,6],pij[,6],pij[,6])
me_6 <- pij_6*(dij-pij)*m1[1,]
me_6 <- apply(me_6,2,mean)
pij_7 <- cbind(pij[,7],pij[,7],pij[,7],pij[,7],pij[,7],pij[,7],pij[,7],pij[,7],pij[,7],pij[,7])
me_7 <- pij_7*(dij-pij)*m1[1,]
me_7 <- apply(me_7,2,mean)
pij_8 <- cbind(pij[,8],pij[,8],pij[,8],pij[,8],pij[,8],pij[,8],pij[,8],pij[,8],pij[,8],pij[,8])
me_8 <- pij_8*(dij-pij)*m1[1,]
me_8 <- apply(me_8,2,mean)
pij_9 <- cbind(pij[,9],pij[,9],pij[,9],pij[,9],pij[,9],pij[,9],pij[,9],pij[,9],pij[,9],pij[,9])
me_9 <- pij_9*(dij-pij)*m1[1,]
me_9 <- apply(me_9,2,mean)
pij_10 <- cbind(pij[,10],pij[,10],pij[,10],pij[,10],pij[,10],pij[,10],pij[,10],pij[,10],pij[,10],pij[,10])
me_10 <- pij_10*(dij-pij)*m1[1,]
me_10 <- apply(me_10,2,mean)
me_all <- rbind(me_1,me_2,me_3,me_4,me_5,me_6,me_7,me_8,me_9,me_10)
View(me_all)


theta2 <- as.vector(optim(c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0), MultLogit, X=income, method = "BFGS")$par)
a2 = c(0,theta2[10:18])
alpha2ij = sapply(a2,rep,4470)
beta2 = c(0,theta2[1:9])
beta2ij = sapply(beta2,rep,4470) 
Vij2 = income*beta2ij + alpha2ij
Pij2 = prop.table(exp(Vij2),1)
beta_i_bar <- Pij2[,1]*beta2ij[,1]+Pij2[,2]*beta2ij[,2]+Pij2[,3]*beta2ij[,3]+Pij2[,4]*beta2ij[,4]+Pij2[,5]*beta2ij[,5]+Pij2[,6]*beta2ij[,6]+Pij2[,7]*beta2ij[,7]+Pij2[,8]*beta2ij[,8]+Pij2[,9]*beta2ij[,9]+Pij2[,10]*beta2ij[,10]
ME2 <- Pij2*(beta2ij - beta_i_bar)
as.matrix(colMeans(ME2))
```



## Exercise 5 IIA
```{r}
mixlogit <- function(theta, X1, X2, Y){
  a = sapply(c(0,theta[11:19]),rep,4470)
  b = c(0,theta[2:10]) 
  b = sapply(b,rep,4470)
  Vij = X1*theta[1] + X2*b + a
  Pij = prop.table(exp(Vij),1)
  Y = -sum(dij*log(Pij))
  return(Y)
}
beta_f <- as.vector(optim(c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0), mixlogit, X1=prt, X2=income, method = "BFGS")$par)
print(beta_f)



xxx = choiceprice
xxx[xxx==1] = NA
xxx1 = na.omit(xxx)

merge2 <- merge(demos[,c(1:2)],xxx1[,c(1:2)], by="hhid")
inc2 <- matrix(rep(merge2[,2],9),ncol=9)

price1 = as.data.frame(xxx1[,2:12])
price3 <- NULL
for(i in 1:9){
  price2 <- price1[,i+1] - price1[,1]
  price3 <- cbind(price3, price2)
}

pd <- subset(choiceprice[,2:12])
pd[pd==1] = NA
pd1 <- na.omit(pd)
dij2n <- ifelse(pd1$choice == "2", 1, 0)
dij3n <- ifelse(pd1$choice == "3", 1, 0)
dij4n <- ifelse(pd1$choice == "4", 1, 0)
dij5n <- ifelse(pd1$choice == "5", 1, 0)
dij6n <- ifelse(pd1$choice == "6", 1, 0)
dij7n <- ifelse(pd1$choice == "7", 1, 0)
dij8n <- ifelse(pd1$choice == "8", 1, 0)
dij9n <- ifelse(pd1$choice == "9", 1, 0)
dij10n <- ifelse(pd1$choice == "10", 1, 0)
dijnew <- cbind(dij2n,dij3n,dij4n,dij5n,dij6n,dij7n,dij8n,dij9n,dij10n)

mixlogit_movecho <- function(theta, X1, X2, Y){
  a = sapply(c(0,theta[10:17]),rep,2699)
  b = c(0,theta[2:9]) 
  b = sapply(b,rep,2699)
  Vij = X1*theta[1] + X2*b + a
  Pij = prop.table(exp(Vij),1)
  Y = -sum(dijnew*log(Pij))
  return(Y)
}

beta_r <- as.vector(optim(c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0), mixlogit_movecho, X1=price3, X2=inc2, method = "BFGS")$par)
print(beta_r)

```